An OpenCL Implementation of Wait-Free Sets

### 2015-06-19 Docs/paper.tex
doc changed

\section{Proof of Correctness}
\begin{algorithm}
  \caption{Deletes all marked nodes beyond {\it startRef}}
  \label{alg:deleteMarkedNode}

  \begin{algorithmic}[1]
    \Require
       \Statex startRef:Reference to the node beyond which all
               marked nodes are to be deleted. 
    \Ensure
       \Statex All marked nodes beyond {\it startRef} are deleted.
       \Statex {\it nextRef} returns next unmarked node beyond {\it startRef}.  
       \Statex Returns {\it true} if all marked nodes are deleted,
               {\it false} if node referenced by {\it startRef} gets deleted.

    \Statex   
    \Function{delMarkedNodes}{startRef, nextRef\&}
      \State doneFlag = false;
      \While{(!doneFlag)}    
        \State [nextRef,startMark] := getUnmarkedRef(startRef);
        \If{(!startMark)}
          \State [nextNextRef,nextMark] := getUnmarkedRef(nextRef);
          \If{(nextMark)}
            \State status = CAS(nodePool[startRef]:next,[nextRef,startMark],[nextNextRef,startMark]);
            \If{(status = true)}
               \State freeNode(nextRef);
            \EndIf
          \Else
            \State
            \Return true;
          \EndIf
        \Else{}
          \State
          \Return false;  
        \EndIf
      \EndWhile        
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
  \caption{Check next node for key}
  \label{alg:check}
  \begin{algorithmic}[5]
    \Function{checkKey}{key, prevRef\&, nodePool}
       \State [nextRef,prevMark] := getPtr(prevRef);
       \If {(prevMark)}
         \State
         \Return {\small NODE\_DELETED}.
       \EndIf
       \If {(nextRef = 0)}
         \State
         \Return {\small KEY\_NOT\_DELETED}.
       \EndIf

       \State [nnRef,nextMark] := getPtr(nextRef);
       \If {(nextMark)}
         \State
         \Return {\small NODE\_NEXT}.
       \EndIf

       \State maxVal := simdMax(nodePool,nextRef);
         \If{maxVal $<$ key}
            \State prevRef := nextRef;
            \Return {\small NODE\_NEXT}.
         \Else
           \If{simdAny(val[lid] = key)}
               \State
               \Return {\small KEY\_FOUND};
           \Else
               \State               
               \Return {\small KEY\_NOT\_FOUND};   
           \EndIf   
         \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Find a key in the set}
  \label{alg:find}
  \begin{algorithmic}[5]
    \Function{find}{key, prevRef\&, nodePool}
       \State lid := simdGetIndex();
       \State startRef := hashFunction(key);
       \State prevRef  := startRef;
       \State doneFlag := false;
       \While{(!doneFlag)}
          \If{(lid = 1)}
             \State delStatus := delMarkedNode(prevRef,nodePool);
          \EndIf
          \State simdBarrier();
          \State delStatus = simdBroadcast(delStatus,1);
          \If{delStatus = true}
             \State keyStatus := checkKey(key, prevRef, nodePool);
             \If{(keyStatus = NODE\_DELETED)}
                \State prevRef := startRef;
             \ElsIf{(keyStatus = KEY\_FOUND)}
                \State 
                \Return true
             \ElsIf{(keyStatus = KEY\_NOT\_FOUND)}
                \State 
                \Return false
             \EndIf
          \Else
             \State prevRef := startRef;   
          \EndIf
       \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Insert a key in the set}
  \label{alg:insert}
  \begin{algorithmic}[5]  
    \Function{insert}{key, nodePool}
      \State doneFlag := false;
      \While{(!doneFlag)}
        \State findStatus := find(key, prevRef\&, nextRef\&, nodePool);
        \If{(findStaus = true)}
          \State
          \Return false;
        \EndIf
        \State minIndex := getMinEmptySlot(nextRef,nodePool);
        \If{minIndex = 0} \Comment No empty slot.
          \State insertStatus := insertNodeAndKey(key, prevRef\&,nextRef\&,nodePool);
        \Else
          \State insertStatus := insertKey(key, minIndex, nextRef\&,nodePool);
        \EndIf
        \If{insertStatus = true}
          \State
          \Return true;
        \EndIf
      \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Find empty slot with minimum index}
  \label{alg:minindex}
  \begin{algorithmic}[5]  
    \Function{getMinEmptySlot}{nextRef, nodePool}
      \State val := nodePool[nextRef]:V[lid];
      \State indicator := (val = EMPTY\_SLOT)?lid:0;
      \If {(simdALL(indicator = 0))}
        \State minIndex := 0;
      \Else
        \State minIndex := simdMin(indicator);
      \EndIf
      \Return minIndex;        
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Insert a node and a key}
  \label{alg:insnodekey}
  \begin{algorithmic}[5]  
    \Function{insertNodeAndKey}{key, prevRef\&, nextRef\&, nodePool}
      \State newNRef := Alloc(nodePool);
      \State newNNRef := Alloc(nodePool);
      \State newNRef:next := [newNNRef,0];
      \If{(nodePool[nextRef]:V[lid] < key)}
        \State nodePool[newNRef]:V[lid] := nodePool[nextRef]:V[lid];
        \State nodePool[newNNRef]:V[lid] := EMPTY\_SLOT;
      \Else
        \State nodePool[newNNRef]:V[lid] := nodePool[nextRef]:V[lid];
        \State nodePool[newNPRef]:V[lid] := EMPTY\_SLOT;
      \EndIf
      \State minIndex := getMinEmptySlot(newNRef,nodePool);
      \State nodePool[newNRef]:V[minIndex] := key;
      \State nodePool[newNNPtr]:next := [nextRef,0];
      \State status := CAS(nodePool[prevRef]:next,[nextRef,0],newNPtr);
      \Return status;
    \EndFunction
  \end{algorithmic}
\end{algorithm}

>>>>

\section{Proof of Correctness}
\begin{algorithm}
  \caption{Deletes all marked nodes that follow a given node}
  \label{alg:deleteMarkedNode}

  \begin{algorithmic}[1]
    \Require
        \Statex startRef:Reference to the starting node. Marked nodes
                that follow this are to be deleted. 
    \Ensure
       \Statex All marked nodes following {\it startRef} are deleted.
       \Statex Returns {\it true} if all marked nodes are deleted,
               {\it false} if node referenced by {\it startRef} gets deleted.

    \Statex   
    \Function{delMarkedNodes}{startRef}
      \State doneFlag = false;
      \While{(!doneFlag)}    
        \State [nextRef,startMark] := getUnmarkedRef(startRef);
        \If{(!startMark)}
          \State [nextNextRef,nextMark] := getUnmarkedRef(nextRef);
          \If{(nextMark)}
            \State status = CAS(startRef:next,[nextRef,startMark],[nextNextRef,startMark]);
            \If{(status = true)}
               \State freeNode(nextRef);
            \EndIf
          \Else
            \State
            \Return true;
          \EndIf
        \Else{}
          \State
          \Return false;  
        \EndIf
      \EndWhile        
    \EndFunction
  \end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}
  \caption{Check next node for key}
  \label{alg:check}
  \begin{algorithmic}[1]
    \Require
       \Statex {\it key} to be checked.
       \Statex {\it prevRef}, reference to node next to which the key is to
               be checked.
    \Ensure
       \Statex Nodes referenced by {\it prevRef} and {\it prevRef:next} are
               unmarked. If they are marked, returns NODE\_DELETED and
               NODE\_NEXT respectively.
       \Statex If all values in {\it prevRef:next} $<$ {\it key}, moves
               {\it prevRef} to {\it prevRef:next}.          
       \Statex If values in {\it prevRef:next} $<$ {\it key}, checks for the key.
               Returns KEY\_FOUND or KEY\_NOT\_FOUND respectively.

    \Statex   
    \Function{checkKey}{key, prevRef\&}
       \State [nextRef,prevMark] := getUnmarkedPtr(prevRef);
       \If {(prevMark)}
         \State
         \Return {\small NODE\_DELETED}.
       \EndIf
       \If {(nextRef = 0)}
         \State
         \Return {\small KEY\_NOT\_FOUND}.
       \EndIf

       \State [nnRef,nextMark] := getUnmarkedPtr(nextRef);
       \If {(nextMark)}
         \State
         \Return {\small NODE\_NEXT}.
       \EndIf

       \State maxVal := simdMAX(nextRef:V[lid]);
       \If{maxVal $<$ key}
           \State prevRef := nextRef;
           \Return {\small NODE\_NEXT}.
       \Else
           \If{simdANY(nextRef:V[lid] = key)}
               \State
               \Return {\small KEY\_FOUND};
           \Else
               \State               
               \Return {\small KEY\_NOT\_FOUND};   
           \EndIf   
         \EndIf
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Find a key in the set}
  \label{alg:find}
  \begin{algorithmic}[1]
    \Require
       \Statex {\it key} to be searched.
    \Ensure
       \Statex {\it prevRef}, such that next node either has key or key could
               be inserted to it.
       \Statex Returns {\it true} if key is found else returns {\it false}.

    \Function{find}{key, prevRef\&}
       \State lid := simdINDEX();
       \State startRef := hashFunction(key);
       \State prevRef  := startRef;
       \State doneFlag := false;
       \While{(!doneFlag)}
          \If{(lid = 1)}
             \State delStatus := delMarkedNode(prevRef);
          \EndIf
          \State simdBARRIER();
          \State delStatus = simdBROADCAST(delStatus,1);
          \If{delStatus = true}
             \State keyStatus := checkKey(key, prevRef);
             \If{(keyStatus = NODE\_DELETED)}
                \State prevRef := startRef;
             \ElsIf{(keyStatus = KEY\_FOUND)}
                \State 
                \Return true
             \ElsIf{(keyStatus = KEY\_NOT\_FOUND)}
                \State 
                \Return false
             \EndIf
          \Else
             \State prevRef := startRef;   
          \EndIf
       \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Insert a key in the set}
  \label{alg:insert}
  \begin{algorithmic}[1]
    \Require
       \Statex {\it key} to be searched.
    \Ensure
       \Statex Returns {\it false} if key is found else inserts the {\it key}.

    \Statex
    \Function{insert}{key}
      \State doneFlag := false;
      \While{(!doneFlag)}
        \State findStatus := find(key, prevRef);
        \If{(findStaus = true)}
          \State
          \Return false;
        \EndIf
        \State [nextRef, prevMark] = getUnmarkedRef(prevRef);
        \If{(prevMark = 0)}
          \State insertStatus = false;
          \State minIndex := getMinEmptySlot(nextRef,nodePool);
          \If{(minIndex = 0)} \Comment No empty slot.
            \State insertStatus := insertNodeAndKey(key, prevRef,nextRef);
          \ElsIf{(minIndex > 0)}
            \State insertStatus := insertKey(key, minIndex, prevRef, nextRef);
          \EndIf
          \If{insertStatus = true}
            \State
            \Return true;
          \EndIf
        \EndIf
      \EndWhile
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Find empty slot with minimum index}
  \label{alg:minindex}
  \begin{algorithmic}[1]
    \Require
      \Statex {\it nextRef} in which minumum index to empty slot is to be
              found.
    \Ensure
      \Statex Returns {\it -1} if {\it nextRef} is deleted, {\it 0} if no
              empty slot is found, or minimum index to empty slot.

    \Function{getMinEmptySlot}{nextRef}
      \State val := nextRef:V[lid];
      \State indicator := (val = EMPTY\_SLOT)?lid:0;
      \If {(simdALL(indicator = 0))}
        \State minIndex := 0;
      \Else
        \State minIndex := simdMIN(indicator);
      \EndIf
      \State [nnRef,nextMark] := getUnmarkedPtr(nextRef);
      \If{(nextMark = 0)}
        \State
        \Return minIndex;
      \Else
        \State
        \Return -1;
      \EndIf  
    \EndFunction
  \end{algorithmic}
\end{algorithm}

\begin{algorithm}
  \caption{Insert a node and a key}
  \label{alg:insnodekey}
  \begin{algorithmic}[1]
    \Require
      \Statex {\it key} to be inserted.
      \Statex {\it prefRef} and {\it nextRef} give window in which new node
              and key are to be inserted.
    \Ensure
       \Statex Returns {\it true} if successful.

    \Function{insertNodeAndKey}{key, prevRef, nextRef}
      \If{(lid = 1)} 
        \State newNRef := newNode();
        \State newNNRef := newNode();
      \EndIf
      \State simdBARRIER();
      \State newNRef  := simdBROADCAST(newNRef,1);
      \State newNNRef := simdBROADCAST(newNRef,1);

      \State newNRef:next := [newNNRef,0];
      \If{(nodePool[nextRef]:V[lid] < key)}
        \State nodePool[newNRef]:V[lid] := nodePool[nextRef]:V[lid];
        \State nodePool[newNNRef]:V[lid] := EMPTY\_SLOT;
      \Else
        \State nodePool[newNNRef]:V[lid] := nodePool[nextRef]:V[lid];
        \State nodePool[newNPRef]:V[lid] := EMPTY\_SLOT;
      \EndIf
      \State minIndex := getMinEmptySlot(newNRef,nodePool);
      \State nodePool[newNRef]:V[minIndex] := key;
      \State nodePool[newNNPtr]:next := [nextRef,0];
      \State status := CAS(nodePool[prevRef]:next,[nextRef,0],newNPtr);
      \Return status;
    \EndFunction
  \end{algorithmic}
\end{algorithm}
